---
title: "dsp_Cow_Lactation_History"
author: "XUHENG LI"
date: "2024-08-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readxl)
library(dplyr)
library(GGally)
library(ggcorrplot)
library(ggplot2)
library(rpart)
library(class)
library(caret)

```


## read dara
```{r}
setwd("C:/Users/13593/Desktop/dsp")
dataset <- read_excel("Cow_Lactation_History.xlsx")
str(dataset)
```

## data preprocess
```{r}
# calculate NA value
na_rate <- sapply(dataset, function(x) sum(is.na(x)) / length(x) * 100)
na_rate_df <- data.frame(Variable = names(na_rate), NA_Percentage = na_rate)
print(na_rate_df)
```
I found that NA_Percentage is very high. And I found that NA values mainly appeared after data with cow number 2250 and before data cow number 1001, so I deleted these rows

```{r}
dataset_sorted <- dataset %>%
  arrange(`Cow Number`)
dataset_filtered <- dataset_sorted %>%
  filter(`Cow Number` >= 1001 & `Cow Number` <= 2219)

head(dataset_filtered)

na_rate1 <- sapply(dataset_filtered, function(x) sum(is.na(x)) / length(x) * 100)

na_rate_df1 <- data.frame(Variable = names(na_rate1), NA_Percentage = na_rate1)

print(na_rate_df1)
```

The NA percent is slightly better, but it's still a lot of NA, so I delete the variables that have too much NA percent, Then I fill the NA values of the remaining columns with the median of that column

```{r}
# select variables
dataset_cleaned <- dataset_filtered %>%
  select(-`Production`, -`Lactose Indication`, -`Lactose indication %`, -`Predicted Production (305 days)`)
str(dataset_cleaned)

# mutate median value
replace_na_with_median <- function(x) {
  if (is.numeric(x)) {
    x[is.na(x)] <- median(x, na.rm = TRUE)
  }
  return(x)
}


dataset_filled <- dataset_cleaned %>%
  mutate(across(everything(), replace_na_with_median))

str(dataset_filled)
summary(dataset_filled)

```

## pre-analysis

```{r}
# Observe the overall distribution of the data
par(mfrow = c(2,5))
boxplot(dataset_filled$`Lac Days`, main = "Lac Days")
boxplot(dataset_filled$`Lactation No`, main = "Lactation No")
boxplot(dataset_filled$`Total Amount Feed`, main = "Total Amount Feed")
boxplot(dataset_filled$`Day Production Avg`, main = "Day Production Avg")
boxplot(dataset_filled$`Milk - Fat kg`, main = "Milk - Fat kg")
boxplot(dataset_filled$`Milk - Protein kg`, main = "Milk - Protein kg")
boxplot(dataset_filled$`Avg Number Of Milkings`, main = "Avg Number Of Milkings")
boxplot(dataset_filled$`Average Milk Speed`, main = "Average Milk Speed")
boxplot(dataset_filled$`Fat indication %`, main = "Fat indication %")
boxplot(dataset_filled$`Protein indication`, main = "Protein indication")

# View data distribution and correlation
ggpairs(dataset_filled[, c(4:13)])

# Whether it is relevant and whether it is significant
corr = cor(dataset_filled[, c(4:13)])
p.mat <- cor_pmat(dataset_filled[, c(4:13)], use = "complete", method = "pearson")
ggcorrplot(corr, hc.order = TRUE, type = "lower", lab = TRUE, p.mat = p.mat, 
    insig = "blank")


p1 <- ggplot(dataset_filled, aes(x = Reproduction)) + geom_bar(fill = "#CF6A1A", colour = "black") + 
    theme_bw()
p1




```
For Day Production Avg, Total Amount Feed (0.7) : This is the variable with the highest correlation with Day Production Avg, indicating that the Total Amount Feed has a significant impact on Day Production.

Avg Number Of Milkings (0.48) : Also has some correlation with Day Production and can be considered as a feature.

Milk-fat kg (0.53) and milk-protein kg (0.22) : Although these have a low direct correlation with daily Milk production, they can be used as characteristic variables for further analysis because they are output-related variables.

Because of the high correlation between fat and protein, I chose only fat for the rest of the modeling

## data analysis
```{r}
set.seed(123)  
train_indices <- sample(seq_len(nrow(dataset_filled)), size = 0.7 * nrow(dataset_filled))
train_data <- dataset_filled[train_indices, ]
test_data <- dataset_filled[-train_indices, ]
```


```{r}
# linear model
linear_model <- lm(`Day Production Avg` ~ `Total Amount Feed` + `Avg Number Of Milkings` + `Milk - Fat kg`, data = train_data)
summary(linear_model)
```
The model showed that Avg Number Of Milkings and Milk-Fat kg had a significant positive effect on daily Production Avg, while Total Amount Feed had a significant negative effect.
The explanatory power of the model (R-squared) was moderate, indicating that the model explained about 52% of the variation in the response variables.
The overall model is significant, and the residual error is small, indicating that the model is well fitted, but there is still room for improvement.

```{r}
predictions_lm <- predict(linear_model, newdata = test_data)

mse_lm <- mean((predictions_lm - test_data$`Day Production Avg`)^2)
print(mse_lm)
```

```{r}
# decision tree
decision_model <- rpart(`Day Production Avg` ~ `Total Amount Feed` + `Avg Number Of Milkings` + `Milk - Fat kg`, data = train_data)
predictions_decision <- predict(decision_model, newdata = test_data)

mse_predictions_decision <- mean((predictions_decision - test_data$`Day Production Avg`)^2)
print(mse_predictions_decision)
```
As we can see, neither of these models is good.

I think the data set is too small for correct data analysis and prediction.


